A meta analysis of tournaments and an evaluation of performance in the Iterated Prisoner's Dilemma.

### 2019-12-29 paper/main.tex
fixes re writes.
This was meant to be 3 different commits but I forgot to commit.
The re writes are the following:

- Conclusion in noisy tourmanets. Explain better that the highly ranked
strategies did so because of their performance in tournaments with p>0.5
- Final conclusion of top rank strategies section. Tweaked the text
a little bit
- Better explanation regarding the clustering of the performances.
The different approaches are now bullet points.

From Figure~\ref{fig:noisy_results}, it is evident that the normalised rank
distributions in noisy environments are more variant with higher medians
compared to standard tournaments. The distributions are bimodal. This indicates
that although the top ranked strategies mainly performed well, there are several
tournaments that they ranked in the bottom half. To gain a better understanding
of the behaviour in noisy tournaments, the \(r\) distributions for the top 6
of Figure~\ref{fig:noisy_results} strategies over the noise probability \(p_n\),
are given in Figure~\ref{fig:effect_of_noise}.

@@

Figure~\ref{fig:effect_of_noise} shows that for \(p_n\) values lower than 0.5
Grumpy, Tit For 2 Tats and Slow Tit For Two Tat perform moderately, and \(e\),
Cycle Hunter and Ricky QLearner perform poorly. At \(p_n=0.5\) all the
distributions become bimodal. This is because with a noise probability of 0.5,
all strategies correspond to a random player. Interestedly, for a \(p_n\) larger
than 0.5 all of the 6 strategies become successful. Note that a value \(p_n=1\)
corresponds to a strategy playing opposite from what it intended to. Thus, it is
demonstrated that the successful strategies is noisy tournaments are sometimes
effective when \(p_n=0.5\) but overall they are very successful whn they are
playing opposite from their original design. If during the data collection a
\(p_n\) strictly less 0.5 was considered then the top ranked strategies would be
different. There are a total of 5661 trials where \(p_n<0.5\) and the top ranked
strategies are given in Table~\ref{table:subset_noise_result}. The median ranks
are lower than before and the top spots are mainly overtaken by Meta strategies
which include NMWE deterministic and NMWE Long Memory. The Meta
strategies~\cite{axelrodproject} create a team of strategies for themselves and
choose to play as a member of their team based on their scores against a given
opponent.

@@

This section presented the winning strategies in a series of IPD tournaments. In
standard tournaments the top spots were dominated by complex strategies that had
been trained using reinforcement learning techniques. In noisy environments,
whether the number of turns was fixed or not, the winning strategies were
deterministic strategies designed to defect if the opponent tricked them more
than a current amount of the times that they had tricked their opponent. However,
if a value of noise strictly less than 0.5 was considered, then the successful
strategies were strategies based on teams. In
probabilistic ending tournaments most of the winning strategies were defecting
strategies and trained finite state automata, designed by the same authors.
These strategies only did better when the probability of the game ending after
each turn was increased.
Finally the performance of all \numberofstrategies
strategies over the \numberofalltournaments tournaments in this manuscript was
assessed on \(\bar{r}\). The top ranked strategies were a mixture of behaviours
that did well in standard tournaments and tournaments with noise, as well as a
few strategies based on teams.

@@

The performances are clustered based on the normalised rank. More specifically,
they are clustered 3 times into 2 different clusters based on  on whether their
normalised rank was in the top 5\%, 25\% or 50\% respectively. A random forest
approach~\cite{breiman2001} is then applied to each performance to predict the cluster to
which it has been assigned to. The random forest method
constructs many individual decision trees and the predictions from all trees are
pooled to make the final prediction. The random forest models are trained on a
training set of 70\% of the tournaments results. The accuracy of each model
based on $R^2$ are given by Table~\ref{table:accuracy_random_forest}. The out of
the bag error~\cite{hastie2005} has also been calculated. The models fit well,
and a high value of both the accuracy measure on the test data and the OOB error
indicate that the model is not over fitting.

The performances have also been clustered based on their normalised rank and
their median score by a \(k-\)means algorithm~\cite{Arthur2007}. The number of
clusters is not deterministically chosen but it is based on the silhouette
coefficients~\cite{Rousseeuw1987}. The chosen cluster for each tournament type,
as well as the accuracy for random forest models are also given in
Table~\ref{table:accuracy_random_forest}.

\begin{table}[!htbp]
    \begin{center}
        \resizebox{.9\textwidth}{!}{
        \begin{tabular}{lccccc}
    \toprule
    Tournament type & Clustering on & Number of clusters & $R^2$ training data &  $R^2$ test data  & $R^2$ OOB score\\
    \midrule
    standard  & top 5\% $r$              & 2  & 0.998831  & 0.987041 & 0.983708 \\
              & top 25\% $r$             & 2  & 0.998643  & 0.978626 & 0.969202 \\
              & top 50\% $r$             & 2  & 0.998417  & 0.985217 & 0.976538 \\
              & $r$  \& normalised score & 2  & 0.998794  & 0.990677 & 0.982959 \\
    \midrule
    noisy     & top 5\% $r$              & 2 & 0.996677  & 0.950572 & 0.935383\\
              & top 25\% $r$             & 2 & 0.996677  & 0.950572 & 0.935383\\
              & top 50\% $r$             & 2 & 0.996677  & 0.950572 & 0.935383\\
              & $r$  \& normalised score & 3 & 0.996677  & 0.950572 & 0.935383\\
    \midrule
    probabilistic ending & top 5\% $r$              & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & top 25\% $r$             & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & top 50\% $r$             & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & $r$  \& normalised score & 2 & 0.999592  & 0.995128 & 0.992819 \\
    \midrule
    noisy probabilistic ending & top 5\% $r$              & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & top 25\% $r$             & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & top 50\% $r$             & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & $r$  \& normalised score & 4 & 0.990490  & 0.813905 & 0.791418\\
    \midrule
    over \numberofalltournaments tournaments & top 5\% $r$               & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & top 25\% $r$              & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & top 50\% $r$              & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & $r$  \& normalised score  & 3 & 0.993396 & 0.913409 & 0.898059 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{Accuracy metrics for random forest models.}
    \label{table:accuracy_random_forest}
\end{table}

The importance that the features of Table~\ref{table:manual_features} had on
each classification task; to which cluster a performance was assigned to based
on the normalised rank, and their normalised rank and median score have been
calculated and are given by Figures~\ref{fig:clustering_importance_standard},
\ref{fig:clustering_importance_noise},~\ref{fig:clustering_importance_probend},
\ref{fig:clustering_importance_probend_noise}
and~\ref{fig:clustering_importance_overall}. These show that the classifiers
stochastic, make use of game and make use of length have no significant effect,
and several of the features that are highted by the importance are inline with
the correlation results. Moreover, the smoothing parameter \(k\) appears to no
have a significant effect either. The most important features based on the
random forest analysis were $C_{r} / C_{median}$ and $C_r / C_{mean}$.

>>>>

From Figure~\ref{fig:noisy_results}, it is evident that the normalised rank
distributions in noisy environments are more variant with higher medians
compared to standard tournaments. The distributions are bimodal. This indicates
that although the top ranked strategies mainly performed well, there are several
tournaments that they ranked in the bottom half. The bimodality of the \(r\)
distributions can be explained by Figure~\ref{fig:noisy_results} which gives the
\(r\) distributions for the top 6 strategies over the noise probability \(p_n\).
Note that for \(p_n=0.5\) a strategy corresponds to a random player and for
\(p_n=1\) a strategy is behaving in the exact opposite way than its design.

@@

From Figure~\ref{fig:effect_of_noise} it is evident that the strategies highly
ranked in noisy environments did so because of their performance in tournaments
with \(p_n>0.5\). In tournaments with a noise probability lower than \(0.5\) the
strategies performed poorly which is why their \(r\) distributions are bimodal.
If during the data collection a \(p_n\) strictly less 0.5 was considered then the top ranked strategies would be
different. There are a total of 5661 trials where \(p_n<0.5\) and the top ranked
strategies are given in Table~\ref{table:subset_noise_result}. The median ranks
are lower than before and the top spots are mainly overtaken by Meta strategies
which include NMWE deterministic and NMWE Long Memory. The Meta
strategies~\cite{axelrodproject} create a team of strategies for themselves and
choose to play as a member of their team based on their scores against a given
opponent.

@@

This section presented the winning strategies in a series of IPD tournaments. In
standard tournaments the top spots were dominated by complex strategies that had
been trained using reinforcement learning techniques. In noisy environments,
whether the number of turns was fixed or not, the winning strategies were
deterministic strategies designed to defect if the opponent tricked them more
than a current amount of the times that they had tricked their opponent. However,
if a value of noise strictly less than 0.5 was considered, then the successful
strategies were strategies based on the behaviour of many strategies. In
probabilistic ending tournaments most of the highly ranked strategies were defecting
strategies and trained finite state automata, all by the authors of~\cite{Ashlock2006, Ashlock2014}.
These strategies ranked high due to their performance in tournaments where the probability of the game ending after
each turn was bigger than 0.1.
Finally the performance of all \numberofstrategies
strategies over the \numberofalltournaments tournaments in this manuscript was
assessed on \(\bar{r}\). The top ranked strategies were a mixture of behaviours
that did well in standard tournaments and tournaments with noise, as well as a
few strategies based on teams.

@@

To further explore the features that contribute to a strategy's success the
performances are divided into clusters based on whether they were successful or
not. A random forest approach~\cite{breiman2001} is then applied to each
performance to predict the cluster to which it has been assigned to. This allows
for the importance of each feature in the classification to be calculated. In
essence, to calculate which are the important features when one is trying to
predict whether a performance was successful or not.

The performances are clustered into successful and unsuccessful clusters based on
4 different approaches. More specifically:

\begin{itemize}
    \item \textbf{Approach 1:} The performances are divided into two clusters based
    on whether their performance was in the top \(5\%\) of their respective tournaments.
    Thus, whether \(r\) was smaller or larger than 0.05.
    \item \textbf{Approach 2:} The performances are divided into two clusters based
    on whether their performance was in the top \(25\%\) of their respective tournaments.
    Thus, whether \(r\) was smaller or larger than 0.25.
    \item \textbf{Approach 3:} The performances are divided into two clusters based
    on whether their performance was in the top \(50\%\) of their respective tournaments.
    Thus, whether \(r\) was smaller or larger than 0.50.
    \item \textbf{Approach 4:} The performances are clustered based on their normalised rank and
    their median score by a \(k-\)means algorithm~\cite{Arthur2007}. The number of
    clusters is not deterministically chosen but it is based on the silhouette
    coefficients~\cite{Rousseeuw1987}.
\end{itemize}

The random forest method constructs many individual decision trees and the
predictions from all trees are pooled to make the final prediction. The random
forest models are trained on a training set of 70\% of the tournaments results.
The accuracy of each model based on $R^2$ and the number of clusters for each
tournament type (because in the case of Approach 4 it is not
deterministically chosen) are given by Table~\ref{table:accuracy_random_forest}.
The out of the bag error (OOB)~\cite{hastie2005} has also been calculated. The
models fit well, and a high value of both the accuracy measure on the test data
and the OOB error indicate that the model is not over fitting.

\begin{table}[!htbp]
    \begin{center}
        \resizebox{.9\textwidth}{!}{
        \begin{tabular}{lccccc}
    \toprule
    Tournament type & Clustering Approach & Number of clusters & $R^2$ training data &  $R^2$ test data  & $R^2$ OOB score\\
    \midrule
    standard  & Approach 1 & 2  & 0.998831  & 0.987041 & 0.983708 \\
              & Approach 2 & 2  & 0.998643  & 0.978626 & 0.969202 \\
              & Approach 3 & 2  & 0.998417  & 0.985217 & 0.976538 \\
              & Approach 4 & 2  & 0.998794  & 0.990677 & 0.982959 \\
    \midrule
    noisy     & Approach 1  & 2 & 0.996677  & 0.950572 & 0.935383\\
              & Approach 2  & 2 & 0.996677  & 0.950572 & 0.935383\\
              & Approach 3  & 2 & 0.996677  & 0.950572 & 0.935383\\
              & Approach 4  & 3 & 0.996677  & 0.950572 & 0.935383\\
    \midrule
    probabilistic ending & Approach 1  & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & Approach 2  & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & Approach 3  & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & Approach 4  & 2 & 0.999592  & 0.995128 & 0.992819 \\
    \midrule
    noisy probabilistic ending & Approach 1  & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & Approach 2  & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & Approach 3  & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & Approach 4  & 4 & 0.990490  & 0.813905 & 0.791418\\
    \midrule
    over \numberofalltournaments tournaments & Approach 1  & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & Approach 2  & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & Approach 3  & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & Approach 4  & 3 & 0.993396 & 0.913409 & 0.898059 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{Accuracy metrics for random forest models.}
    \label{table:accuracy_random_forest}
\end{table}

The importance that the features of Table~\ref{table:manual_features} had on
each random forest model while the perfomances were clustered based on the
different approaches have been
calculated and are given by Figures~\ref{fig:clustering_importance_standard},
\ref{fig:clustering_importance_noise},~\ref{fig:clustering_importance_probend},
\ref{fig:clustering_importance_probend_noise}
and~\ref{fig:clustering_importance_overall}. These show that the classifiers
stochastic, make use of game and make use of length have no significant effect,
and several of the features that are highted by the importance are inline with
the correlation results. Moreover, the smoothing parameter \(k\) appears to no
have a significant effect either. The most important features based on the
random forest analysis were $C_{r} / C_{median}$ and $C_r / C_{mean}$.
